---
Title: "Reproducible Research - Peer Assignment1"
Name: "Nithin V"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading and Reprocessing the data

### - Load the data
```{r, echo=TRUE}
library(ggplot2)
library(stats)
read <- read.csv(file = "activity.csv")
View(read)
summary(read)
dim(read)
```

# What is mean total number of steps taken per day?

### - Sum steps by date
```{r}
steps_date <- aggregate(steps ~ date, read, sum)
```

### - Create Histogram
```{r}
hist(steps_date$steps, main = "Total Number of Steps Each Day", col = "green", xlab = "Number of Steps")
```

### - Calculate Mean and Median
```{r}
Mean <- mean(steps_date$steps)
Median <- median(steps_date$steps)
```

* Mean: `r Mean` 
* Median: `r Median`


# What is the Average Daily Activity Pattern

### - Calculate the average steps for each interval
```{r}
steps_interval <- aggregate(steps ~ interval, read, mean)
```

### - Plot the Average Number of Steps per day with respect to interval
```{r}
plot(steps_interval$interval, steps_interval$steps, type = "l", main = "Average Steps / Day by Interval", xlab = "Interval", ylab = "Number of Steps")
```

### - Interval with most average steps
```{r}
max_int <- steps_interval[which.max(steps_interval$steps), 1]
```


* Maximam 5 min Interval: `r max_int` 

# Imputing missing values

### - Calculate and report the total number of missing values in the dataset
```{r}
any(is.na(read$steps))
count <- sum(is.na(read$steps))
```

* Missing Values: `r count`

### - Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc

### - Create a new dataset that is equal to the original dataset but with the missing data filled in
```{r}
fil <- function(steps, interval) {
  filler <- NA
  if (!is.na(steps))
    filler <- c(steps)
  else
    filler <- (steps_interval[steps_interval$interval==interval, "steps"])
  return(filler)
}
filler.new <- read
filler.new$steps <- mapply(fil, filler.new$steps, filler.new$interval)
```

### - Histogram showing the number of steps taken each day

```{r}
Tsteps <- tapply(filler.new$steps, filler.new$date, FUN = sum)
qplot(Tsteps, binwidth = 2000, xlab = "Total number of steps taken each day", main = "Histogram showing the number of steps taken each day")
Mean1 <- mean(Tsteps)
Median1 <- median(Tsteps)
```

* Mean: `r Mean1`
* Median: `r Median1`

# Are there differences in activity patterns between weekdays and weekends?

### - Create a new factor variable in the dataset with two levels indicating whether a given date is a weekday or weekend day
```{r}
filler.new$daytype <- ifelse(as.POSIXlt(filler.new$date)$wday %in% c(0,6), 'weekend', 'weekday')
```
### - Make a panel plot containing a time series plot
```{r}
final <- aggregate(steps ~ interval + daytype, filler.new, mean)
ggplot(final, aes(interval, steps)) + geom_line() + facet_grid(daytype ~ .)
```
